---
title: "a3_task4_karla_garibay"
author: "Karla Garibay Garcia"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```

## Summary: In this document, I will be # TODO


## Attach packages

Here, I attach tidyverse, janitor, here, kableExtra, and lubridate packages in the setup code chunk. 


## Read in data

Here, I read in the urchins.csv data from the data folder. 

```{r}
urchins.csv <- read.csv(here("data", "urchins.csv"))
```

## Clean up data

Here, I use clean_names() to switch all column names into lower_snake_case. 

```{r}
urchins_clean <- urchins.csv %>% clean_names()
```

## Data wrangling

Here, I create a subset of the urchins data called naples_urchins that: 

- Only retains observations from Naples Reef from January 2010

```{r, warning = FALSE, message = FALSE}
naples_urchins <- urchins_clean %>% 
  filter(site == "NAPL") %>%
  mutate(date_new = mdy(date)) %>% 
  mutate(month = month(date_new)) %>% 
  mutate(year = year(date_new)) %>% 
  filter(month == 1) %>% 
  filter(year == 2010)
```

## Exploratory data visualization 

Here, I use a faceted histrogram and boxplot to compare the distribution of the Annual and Control treatments to assess normality. 

```{r}
ggplot(data = naples_urchins, 
       aes(x = size)) + 
  geom_histogram(bins = 15) +
  labs(x = "Size(cm)", 
       y = "Count") +
  facet_wrap(~treatment)
```

```{r}
ggplot(data = naples_urchins, 
       aes(x = treatment, 
           y = size)) +
  geom_boxplot() +
  labs(x = "Treatment type", 
       y = "Size(cm)")
```

### Takeaways from exploratory data visualization # TODO


## Descriptive statistics table

Here, I create a summary table that contains the mean, median, standard deviation, and sample size statistics for Annual and Control treatments. 

```{r, warning = FALSE, message=FALSE}
tx_stats <- naples_urchins %>% 
  group_by(treatment) %>% 
  summarise(mean_size = mean(size), 
            median_size = median(size),
            sd_size = sd(size), 
            n = n())

tx_stats %>% 
  kbl() %>% 
  kable_styling()
```

## Two-sample t-test

Here, I perform a two-sample t-test to compare means between the Annual and Control treatment groups by:

- Filtering to only contain observations for annual and control samples from naples_urchins
- Using the pull() function to only pull the values in the size column and store them as vectors for both samples

```{r}
annual_sample <- naples_urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- naples_urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x = annual_sample, control_sample)
```

### Takeaways from two-sample t-test #TODO



## P-value meaning #TODO





